---
title: "class15"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## RNA Seq Analysis

#Install DESeq2 Bioconductor Package
```{r}
#source("http://bioconductor.org/biocLite.R")
#biocLite()
# For this class, you'll also need DESeq2:
#biocLite("DESeq2")

library(BiocInstaller)
```

#Read countData and colData (metadata) files that we will use with DESeq
```{r}
counts <- read.csv("data/airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("data/airway_metadata.csv", stringsAsFactors = FALSE)
```


#Examine metadata file to find control and treated columns (cell lines)
```{r}
control.inds <- metadata$dex == "control"
control <- metadata[control.inds,]
#control
```

```{r}
control.mean <- rowSums(counts[,control$id])/nrow(control)
names(control.mean) <- counts$ensgene
#control.mean
```

#Do the same for the treated column

```{r}
treated.inds <- metadata$dex == "treated"
treated <- metadata[treated.inds,]
#treated
treated.mean <- rowSums(counts[,treated$id])/nrow(treated)
names(treated.mean) <- counts$ensgene
#treated.mean
```


#Combine meancount for bookkeeping purposes
```{r}
meancounts <- data.frame(control.mean,treated.mean)
colSums(meancounts)
```


#Plot mean of the treated samples against the mean of the control samples with log transformation
```{r}
plot(meancounts$control.mean, meancounts$treated.mean, xlab = "Control", ylab = "Treated", log = "xy")
```

# Log 2 fold change
```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```

# Filter genes with zero expression
```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```

#Filter the dataset for upregulated and downregulated (based on fold change +2/-2)
```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```

#Number of genes upregulated and downregulated
```{r}
sum(up.ind)
sum(down.ind)
```

Annotate the genes
```{r}
anno <- read.csv("data/annotables_grch38.csv")
head(anno)
```

#Merge datasets anno and mycounts
```{r}
mycounts.anno <- merge(mycounts,anno,by.x = 0, by.y = "ensgene")
View(mycounts.anno)
```


#Bioconductor’s annotation packages help with mapping various ID schemes to each other. Here we load the AnnotationDbi package and the annotation package org.Hs.eg.db.
```{r}
#biocLite("AnnotationDbi")
#biocLite("org.Hs.eg.db")
library("AnnotationDbi")
library("org.Hs.eg.db")
```


#List of all available key types
```{r}
columns(org.Hs.eg.db)
```

```{r}
mycounts$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")
#head(mycounts)
mycounts$entrez <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="ENTREZID",
                     keytype="ENSEMBL",
                     multiVals="first")
#head(mycounts)
mycounts$uniprot <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="UNIPROT",
                     keytype="ENSEMBL",
                     multiVals="first")
head(mycounts)
```

```{r}
head(mycounts[up.ind,])
head(mycounts[down.ind,])
```

#Load DESeq2 package
```{r}
library(DESeq2)
```

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```

```{r}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
res
```

#Print a summary of our DESeq results
```{r}
summary(res)
```


#Print a summary of our DESeq results with a p-value of 0.05
```{r}
res05 = results(dds, alpha = 0.05)
summary(res05)
```

#Print a summary of our DESeq results with a p-value of 0.01
```{r}
res01 = results(dds, alpha = 0.01)
summary(res01)
View(res01)
```

#Annotate results stored in object res01

```{r}
res01$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(res01),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")
head(res01)
```

#Save results to csv file
```{r}
write.csv(res01, "signif01_results.csv")
```

#Focus on thegene CRISPLD2
```{r}
i <- grep("CRISPLD2", res01$symbol)
res01[i,]
```

#Use the DESeq2 package function plotCounts()
```{r}
plotCounts(dds,gene="ENSG00000103196", intgroup="dex")
```

```{r}
d <- plotCounts(dds,gene="ENSG00000103196", intgroup="dex", returnData=TRUE)
d
```

```{r}
boxplot(count ~ dex , data=d)
```

#Volcano plots
```{r}
plot(res01$log2FoldChange, -log(res01$padj), xlab = "Fold Change", ylab = "p-value")
```

#Colored version of plot highlighting genes with high fold change and small p-value
```{r}
mycols <- rep("black", nrow(res01))
mycols[(res01$log2FoldChange < -2) & (res01$padj < 0.01)] <- "red"
mycols[(res01$log2FoldChange > 2) & (res01$padj < 0.01)] <- "green"

plot(res01$log2FoldChange, -log(res01$padj), xlab = "Fold Change", ylab = "p-value", col = mycols)
```

