---
title: "Class11: Structural Bioinformatics"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lets Look at the table
```{r}
prot_structure <- read.csv("C:/Users/chinm/Desktop/BIMM 143/bimm143_github/class11/Data Export Summary.csv", row.names = 1)
library(knitr)
kable(prot_structure)
```

# What percentage of structures solved by X-Ray and Electron Microscopy.

```{r}
# Total number of entries
nstru <- sum(prot_structure$Total)

# Percent of structures solved
solved_structures <- round(prot_structure$Total/nstru * 100, 2)

# Add to table
prot_structure$Percent <- solved_structures

#Print percent of solved protein structures
kable(prot_structure)

```

There are `r prot_structure["X-Ray", "Percent"]` % of X-Ray Structures 

There are `r prot_structure["Electron Microscopy", "Percent"]` % of Electron Microscopy Strucutres

# From the website what proportion of structures are protein?

```{r}
#Proportion of structures that are proteins
percent_proteins <- round(sum(prot_structure$Proteins)/nstru *100, 2)

```

The proportion of structures that are proteins is `r percent_proteins` %

# Section 3

```{r}
library(bio3d)
#Reading PDB file data into R
pdb <- read.pdb("1hsg")
print(pdb)

#find the attributes
attributes(pdb)

head(pdb$atom)

# Print a subset of $atom data for the first two atoms
pdb$atom[1:2, c("eleno", "elety", "x","y","z")]

# Note that individual $atom records can also be accessed like this
pdb$atom$elety[1:2]

# Which allows us to do the following
plot.bio3d(pdb$atom$b[pdb$calpha], sse=pdb, typ="l", ylab="B-factor")

str(pdb$atom)

# Print a summary of the coordinate data in $xyz
pdb$xyz

# Examine the row and column dimensions
dim(pdb$xyz)

# Print coordinates for the first two atom
pdb$xyz[ 1, atom2xyz(1:2) ]

# Select all C-alpha atoms (return their indices)
ca.inds <- atom.select(pdb, "calpha")
ca.inds

# Print details of the first few selected atoms
head( pdb$atom[ca.inds$atom, ] )

# And selected xyz coordinates
head( pdb$xyz[, ca.inds$xyz] )

# Select chain A
a.inds <- atom.select(pdb, chain="A")
# Select C-alphas of chain A
ca.inds <- atom.select(pdb, "calpha", chain="A")
# We can combine multiple selection criteria to return their intersection
cab.inds <- atom.select(pdb, elety=c("CA","CB"), chain="A",
resno=10:20)

```

# Section 4 ALigning multiple structures

```{r}
# Download some example PDB files
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A")
files <- get.pdb(ids, split = TRUE)

# Extract and align the chains we are interested in
pdbs <- pdbaln(files, fit = TRUE)
# Print to screen a summary of the 'pdbs' object
pdbs

# find invariant core
core <- core.find(pdbs)
# superimpose all structures to core
pdbs$xyz = pdbfit(pdbs, core)
# Perform PCA
pc.xray <- pca(pdbs)

plot(pc.xray)

# Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")

```

