---
title: "Class 8"
author: "Chinmay Kalluraya"
date: "October 25, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means clustering

#### Our first example with **k-means()** function

```{r, include=FALSE}
# Example plot to see how Rmarkdown works
plot(1:10, typ = "l", echo = FALSE)
```

```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```
# Use the kmeans() function setting k to 2 and nstart=20
```{r}
cl1 <- kmeans(x, center = 2, nstart = 20)
```

# Inspect/print the results

# Q. How many points are in each cluster?
```{r}
cl1$size
```

# Q. What ‘component’ of your result object details
# - cluster size? 
```{r}
cl1$size
```

# - cluster assignment/membership?
```{r}
cl1$cluster
```

# - cluster center?
```{r}
cl1$center
```

# Plot x colored by the kmeans cluster assignment and
# add cluster centers as blue points
```{r}
palette(c("red", "green"))
plot (x, col = cl1$cluster)
points(cl1$centers, col = "blue", pch=20, cex = 2)
```

# Q. Repeat for k=3, which has the lower tot.withinss? 
```{r}
cl2 <- kmeans(x, center = 2, nstart = 20)
cl3 <- kmeans(x, center = 3, nstart = 20)
cl4 <- kmeans(x, center = 4, nstart = 20)
cl5 <- kmeans(x, center = 5, nstart = 20)

scree = (c(cl2$tot.withinss, cl3$tot.withinss, cl4$tot.withinss, cl5$tot.withinss))
plot(scree, typ="b", main = "scree")
```


## Heirarchial Clustering in R

#### Let's try out the **hclust()** function for Heirarchial clustering in R. This function needs a distance matrix as input.

```{r}
d <- dist (x)
hc <- hclust(d)
plot(hc)
abline (h=8, col = "red")
cutree(hc, h=8)
```


# Step 1. Generate some example data for clustering
```{r}
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
```

# Step 2. Plot the data without clustering
```{r}
plot(x)
```

# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
```{r}
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
palette(c("red", "blue", "black"))
plot(x, col=col, pch = 20)
```

# Q. Use the dist(), hclust(), plot() and cutree()
# functions to return 2 and 3 clusters
```{r}
d <- dist (x)
hc <- hclust(d)
plot(hc)
clus3 <- cutree(hc, k=3)
clus2 <- cutree(hc, k=2)
plot(x, col = clus3, pch = 20)
plot(x, col = clus2, pch = 20)
```


# Q. How does this compare to your known 'col' groups?
```{r}
table(clus3, col)
table(clus2, col)
```


## Prinipal component analysis

```{r}

## Initialize a blank 100 row by 10 column matrix
mydata <- matrix(nrow=100, ncol=10)

## Lets label the rows gene1, gene2 etc. to gene100
rownames(mydata) <- paste("gene", 1:100, sep="") 

## Lets label the first 5 columns wt1, wt2, wt3, wt4 and wt5
## and the last 5 ko1, ko2 etc. to ko5 (for "knock-out")
colnames(mydata) <- c( paste("wt", 1:5, sep=""), 
                       paste("ko", 1:5, sep="") ) 

## Fill in some fake read counts
for(i in 1:nrow(mydata)) {
 wt.values <- rpois(5, lambda=sample(x=10:1000, size=1))
 ko.values <- rpois(5, lambda=sample(x=10:1000, size=1))

 mydata[i,] <- c(wt.values, ko.values)
}
head(mydata)

## Try to find structure in this data wth prcomp

## lets do PCA
pca <- prcomp(t(mydata), scale=TRUE)
## See what is returned by the prcomp() function
attributes(pca) 

## A basic P1 vs PC2 2-D Plot

plot(pca$x[,1], pca$x[,2])

## Precent variance is often more informative to look at
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1) 

## Make Scree Plot
barplot(pca.var.per, main="Scree Plot",
 xlab="Principal Component", ylab="Percent Variation")

## A vector of colors for wt and ko samples
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
 xlab=paste0("PC1 (", pca.var.per[1], "%)"),
 ylab=paste0("PC2 (", pca.var.per[2], "%)")) 

## Lets focus on PC1 as it accounts for > 90% of variance
loading_scores <- pca$rotation[,1]
summary(loading_scores) 

## We are interested in the magnitudes of both plus
## and minus contributing genes
gene_scores <- abs(loading_scores) 

## Sort by magnitudes from high to low
gene_score_ranked <- sort(gene_scores, decreasing=TRUE) 

## Find the names of the top 5 genes
top_5_genes <- names(gene_score_ranked[1:5])
## Show the scores (with +/- sign)
pca$rotation[top_5_genes,1]

## UK Food Data

x <- read.csv("UK_foods.csv")
dim(x)
rownames(x) <- x[,1]
x <- x[,-1]
head(x)
dim(x)
knitr::kable(x, caption="The full UK foods data table")
heatmap(as.matrix(x))
pca <- prcomp( t(x) )
summary(pca)
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x))
v <- round( pca$sdev^2/sum(pca$sdev^2) * 100 )
v
z <- summary(pca)
z$importance
barplot(v, xlab="Principal Component", ylab="Percent Variation")


```

